### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/empirical.tex
Empirical results update

\section{Empirical}
In this section, we conduct experiments to study the feature selection performances of several well known methods such as Symmetrical Uncertainty Rank, Gain Ratio Rank, Information Gain Rank, Correlation Based Rank, Conditional Entropy Rank \cite{guyon_jmlr03}, Correlation-based Feature Subset Selection \cite{Hall1998},  Reliff \cite{Robnik-Sikonja2003} , mRMR \cite{peng2005} and our proposed methods: High Precision Expectation, High Precision Log likelihood, High Recall Expectation and High Recall Log likelihood. The first six methods are variable rank methods and the last six are subset rank methods.

In these experimental study, we evaluated each feature selection method in binary classification problems such as Breast Cancer, Pima Diabetes, Heart Statlog, Spect and Vote. These data sets can be downloaded from UCI machine learning repository \cite{Bache+Lichman:2013}. In addition, we used three different classifiers to solve each of the binary classifier problem. Logistic Regression, SVM Linear and Naive Bayes were used in the experiments and the first two are implemented in LibLinear library \cite{REF08a} and the last one is implemented in Weka software \cite{weka}.

To evaluate the feature selection performances we reported classification metrics performing 10-fold cross-validation from one until the max number of features selected for each classifier solving each of the binary classification problem. Specifically, we reported Accuracy, Precision, Recall and F-Score for the true label class, that could be identified by the context of the problem and usually be the class with least samples in the data set.

>>>>

\section{Empirical}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[tbp!]
\centering
\begin{tabular}{|l|r|r|r|r|l|} \hline
Dataset & \# Features & \# Data & \# Features/\# Data & \% True Labels & Feature Type\\ \hline \hline 
Breast Cancer	& 9	& 699	& 0.013	& 34 \%	& All Numeric\\ \hline
Diabetes        & 8	& 768	& 0.010	& 35 \%	& All Numeric\\ \hline
Heart Statlog	& 13	& 270	& 0.048	& 44 \%	& All Numeric\\ \hline
Spect	        & 22	& 80	& 0.275 & 33 \%	& Categorical\\ \hline
Vote	        & 16	& 435	& 0.037	& 39 \%	& Categorical\\ \hline
Newsgroup	& 500	& 1963	& 0.255	& 49 \%	& All Binary\\ \hline
Horse Colic	& 22	& 368	& 0.060	& 37 \%	& Categorical (16) and numeric (6)\\ \hline
Credit-American	& 15	& 690	& 0.022	& 44 \%	& Categorical(9) and numeric (6)\\ \hline
Credit-German	& 20	& 1000	& 0.020 & 30 \%	& Categorical (13) and Numeric (7)\\ \hline
Hepatitis	& 19	& 155	& 0.123	& 21 \%	& Categorical (13) and Numeric (6)\\ \hline
Ionosphere	& 34	& 351	& 0.097	& 36 \%	& All Numeric\\ \hline
%KR-vs-KP	& 36	& 3196	& 0.011	& 48 \%	& All Categorical\\ \hline
%Labor	        & 16	& 57	& 0.281	& 35 \%	& Categorical (8) and Numeric (8)\\ \hline
%Mushroom	& 22	& 8124	& 0.003	& 48 \%	& All Categorical\\ \hline
%Sick	        & 29	& 3772	& 0.008	&  6 \%	& Categorical (23) and Numeric (6)\\ \hline
%Sonar	        & 60	& 208	& 0.288 & 47 \%	& All Numeric\\ \hline
\end{tabular}
\caption{\footnotesize Statistics of the various datasets evaluated in this work.}
\label{table:stats}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank (SUR), Gain Ratio Rank (GRR), Information Gain Rank
(IGR), Correlation Based Rank (CBR), Conditional Entropy Rank (CER)~\cite{guyon_jmlr03}, 
Correlation-based Feature Subset Selection (CFS)~\cite{Hall1998}, 
Reliff (R)~\cite{Robnik-Sikonja2003} , mRMR~\cite{peng2005} and our proposed 
methods: High Recall Expectation (HRE) and High Recall Log likelihood (HRL), 
High Precision Expectation (HPE), High Precision Log likelihood (HPL). 
The first five methods are variable rank methods and the last seven
are subset rank methods. The first five methods are variable rank methods that select variables by ranking them with some metric and the last seven are subset rank methods that assess subsets of variables and consider previous selections to decide the next selection.

{\bf TODO Rodrigo: please update acronyms in plots with those given above,
use MRMR for MRMR... sorry, does not seem sensible to shorten this to 3 letters.}

% TODO: Scott... any interesting things to point out?  
%       Class imbalance / feature-data ratios?

In this experimental study, we evaluated each feature selection
method on a variety of binary classification problems 
from the UCI machine learning repository \cite{Bache+Lichman:2013}.
These datasets along with their properties are outlined in 
Table~\ref{table:stats}.

In addition, we used three different classifiers to solve each of the
binary classifier problem. Logistic Regression (LR), SVM Linear (SVM) and Naive
Bayes (NB) were used in the experiments and the first two are implemented
in the LibLinear library~\cite{REF08a} and the last one is implemented in
the Weka software~\cite{weka}.

To evaluate feature selection algorithm performance, we perform 
10-fold nested cross-validation (nesting for tuning hyperparameters of each classification
algorithm) and evaluate accuracy, precision, and recall at each stage of
feature selection from the first selected feature through to the maximum
(either 50 or the number of features in the dataset, whichever is smaller).
%and F-Score for the true label class, that
%could be identified by the context of the problem and usually be the
%class with least samples in the data set.

In our evaluation, we wish to answer the following questions:
Do different feature selection methods perform better with
each classifier or on each dataset in comparison to others?
How reliably do the different feature selection methods perform overall
in terms of their performance distribution?
%\item Do any dataset features seem to correlate with algorithm performance?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/BarGraphs/dataSetXFeatureSelection_avgCl/Accuracy.pdf}\\
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/BarGraphs/dataSetXFeatureSelection_avgCl/Recall.pdf}
\caption{\footnotesize Performance of various feature selection algorithms per dataset, averaged across classifier type and each stage of feature selection.}
\label{fig:perf_vs_dataset}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/BarGraphs/bargraph4/Accuracy.pdf}\\
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/BarGraphs/bargraph4/Recall.pdf}
\caption{\footnotesize Performance of various feature selection algorithms per classifier, averaged across dataset and each stage of feature selection.}
\label{fig:perf_vs_classifier}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\centering
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/LineGraphs/FullResultsWithoutNoutOfK/Spect Logistic Regression Recall.pdf}
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/LineGraphs/FullResultsWithoutNoutOfK/Spect Naive Bayes Recall.pdf}\\
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/LineGraphs/FullResultsWithoutNoutOfK/Newsgroup Logistic Regression Recall.pdf}
\includegraphics[width=0.35\textwidth]{../ResultsPlot/OthersDataSets/LineGraphs/FullResultsWithoutNoutOfK/Newsgroup Naive Bayes Recall.pdf}
\caption{\footnotesize Performance distribution of various feature selection algorithms with samples taken for each dataset, classifier type, and stage of feature selection.}
\label{fig:perf_vs_fs_alg}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To answer these questions, we first start with
Figure~\ref{fig:perf_vs_dataset}, where we evaluate the performance of
various feature selection algorithms per dataset, averaged across
classifier type and each stage of feature selection.  We observe...
% TODO Scott

In Figure~\ref{fig:perf_vs_classifier}, we examine how well our 
novel feature selection algorithms HRE, HRL, HPE, HPL 
perform vs. classifier, noting that overall performance is best
for Naive Bayes followed by Logistic Regression and noticeably worse for
the SVM --- it seems simply that the probabilistic nature of our feature selection
dovetail best with classification methods that are themselves probabilistic,
and overall best with Naive Bayes which makes the same independence assumptions.
{\bf TODO Rodrigo: can we get bargraphs like those in FeatureSelection classifier,
except only showing HRE, HRL, HPE, HPL *and* averaging over all datasets?
So major group={HRE, HRL, HPE, HPL}, minor group={SVM,LR,NB} and one graph
for each of A/P/R/F.  Also the filename should not have spaces.}

Next we examine Figure~\ref{fig:perf_vs_fs_alg}, where we evaluate the
performance distribution of various feature selection algorithms with
samples taken for each dataset, classifier type, and stage of feature
selection.  Here we see Y.  We observe...

### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/intro.tex
Empirical results update

\COMMENT
- Feature selection motivations
** too little data (e.g., bioinformatics, or text with few labels)
** features costly to evaluate
** explainability
** minimize feature elicitation for future data collection (questionnaires, medical tests -- blood tests)
** computation / Big Data
** mobile devices (RAM, computation)
** linear speed (LibLinear 1000x faster than LibSVM)
** poly kernels without kernelized approach without loss of evaluation efficiency
** need to argue SVM/logistic L1 not always best... not always logistic regression/SVM problem, also eval efficient, but training not, so 
inefficient for large-scale or resource-constrained training... show empirical better than L1, also more control over amount of feature 
selection (need a tool that implements full regularization path -- constrains optimization method)

>>>>

Naturally do not want to optimize for precision and recall directly (all false trivially optimizes precision, while all true trivially optimizes accuracy).  But when selecting features for a classifier intended to maximize accuracy, we may have a secondary preference for features that improve precision (reduce false positive rate) or improve recall (reduce false negative rate).  To this end, we remark that no previous work seems to have examined feature selection in the context of the precision-recall trade-off/

% Part of the reason is that previous feature selection has not been specific to classification whereas precision/recall are

% Consider the feature set jointly, weighted voting model

% We get a heavier weighting of effect of data... a feature that pushes more things to their
% class improves recall... so summation is reason for recall.  Each data gets it's own feature vote.

% We get best performance with prob. classifiers, due to alignment of assumptions

% We get best results with conjunction on high ratio of features to data
% - conjunction forces agreement and reduces noise
% - summation encourages coverage of the data
% ... this works best when lot's of features, let data vote, but get probabilistic agreement

% Show anecdotal results for Newsgroup and Spect logistic... we're best across both

in well-known contingency table

\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Label T} & \multicolumn{1}{l}{Label F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}

accuracy = tp + tn / tp + fp + fn + tn

precision = tp / tp + fp

recall = tp / tp + fn

can switch in columns, but typically predicting more false hurts recall and more true hurts precision

So precision and recall become secondary objectives after accuracy... provides flexibility in feature selection that has not been explicitly considered.

% Yet on other hand certain methods must have done better/wose b/c of implicit assumptions that we should analyze later.

\COMMENT
key insight: literature suggests redundancy is bad idea, but in fact, it is critical for precision (implies agreement)

how to strongly justify the model and objectives we use?  voting?

Good discussion: http://jmlr.org/papers/volume13/brown12a/brown12a.pdf

- Feature selection motivations
** too little data (e.g., bioinformatics, or text with few labels)
** features costly to evaluate
** explainability
** minimize feature elicitation for future data collection (questionnaires, medical tests -- blood tests)
** computation / Big Data
** mobile devices (RAM, computation)
** linear speed (LibLinear 1000x faster than LibSVM)
** poly kernels without kernelized approach without loss of evaluation efficiency
** need to argue SVM/logistic L1 not always best... not always logistic regression/SVM problem, also eval efficient, but training not, so 
inefficient for large-scale or resource-constrained training... show empirical better than L1, also more control over amount of feature 
selection (need a tool that implements full regularization path -- constrains optimization method)
